---
title: "CardioRVAR Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CardioRVAR Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This tutorial is intended to serve as a general guide for the CardioRVAR package.
To start the package, use the following command in R:

```{r setup}
library(CardioRVAR)
```

With this, the functions available in the package can now be used. We will use the example
data known as Cardiovascular:

```{r Get data, eval = TRUE}
data("Cardiovascular")
Cardiovascular <- ResampleData(Cardiovascular, f = 4)
Data <- cbind(SBP = Cardiovascular$SBP, IBI = Cardiovascular$RR)
```


To examine the recordings, use the following commands:

```{r Plot recordings, eval = TRUE, fig.align = "center", fig.width = 7}
par(mfrow = c(1,2))
plot(Cardiovascular$Time, Data[,"IBI"], xlab = "Time", ylab = "IBI", type = "l")
plot(Cardiovascular$Time, Data[, "SBP"], xlab = "Time", ylab = "SBP", type = "l")
```

## PART 1: CHECK THAT RECORDS ARE STATIONARY


Our registers are in the "control" and "tilt" objects.
These records belong to RCC and correspond to the interval
pre-tilt and tilt respectively. Each object is an array of
N x M, where N is the number of observations, and M is the number of variables
(2 in this case).

We will use the tilt registers. The trend of these records
and its mean have already been removed, so they should be stationary.
To check this, we use the CheckStationarity function:


```{r Check stationarity, eval = TRUE}
CheckStationarity(Data, verbose = TRUE)

# If we do not supply the "verbose" argument, we should get a
# value of type TRUE/FALSE (in this case, TRUE):

CheckStationarity(Data)

```
As there is a breakpoint, the data is not stationary. We can make the data stationary by detrending it:
```{r Check stationarity again, eval = TRUE}
Detrended_Data <- Data

Detrended_Data[,"IBI"] <- DetrendWithCutoff(Data[,"IBI"])
Detrended_Data[,"SBP"] <- DetrendWithCutoff(Data[,"SBP"])

CheckStationarity(Detrended_Data, warnings = FALSE)
```



Now, we will separate the data into two stationary subsets, Pre and Post. We will also detrend this data:

```{r Check stationarity again 2, eval = TRUE}
Pre <- Data[1:400,]
Post <- Data[500:928,]

Pre[,"IBI"] <- DetrendWithCutoff(Pre[,"IBI"])
Pre[,"SBP"] <- DetrendWithCutoff(Pre[,"SBP"])
Post[,"IBI"] <- DetrendWithCutoff(Post[,"IBI"])
Post[,"SBP"] <- DetrendWithCutoff(Post[,"SBP"])

CheckStationarity(Pre, warnings = FALSE)
CheckStationarity(Post, warnings = FALSE)

```



## PART 2: ESTIMATE VAR MODEL (PART 1)


Once the records have been verified to be
stationary, you can go to the next step, which
consists of estimating a VAR model. It can be done in two ways.
For the first one, we'll use the tilt object, and the VAR function of the package
vars:

```{r Model data in time domain, eval = TRUE}
model_pre <- vars::VAR(Pre, type = "none", lag.max = 22, ic = "AIC")
model_post <- vars::VAR(Post, type = "none", lag.max = 22, ic = "AIC")

c(model_pre$p, model_post$p)
```

With this we have estimated a VAR model of order p, or VAR(p). This
model is of order 12, that is, VAR(12). To check it, simply
we wrote



Now the next step is to validate the model. To do this, we use the functions
DiagnoseResiduals and DiagnoseStability. For the model to be valid, it must be
stable and its residuals must be white noise processes:

```{r Diagnose models, eval = TRUE}
DiagnoseResiduals(model_pre, verbose = TRUE)
DiagnoseStability(model_pre, verbose = TRUE)

DiagnoseResiduals(model_post, verbose = TRUE)
DiagnoseStability(model_post, verbose = TRUE)
```

The results should confirm that the model is valid

## PART 3: ESTIMATE VAR MODEL (PART 2)

The previous steps can be performed automatically using the following function:

```{r Generate full model automatically, eval = TRUE}
model_pre <- EstimateVAR(Pre)
model_post <- EstimateVAR(Post)

c(model_pre$p, model_post$p)
```



## PART 4: Estimating Transfer Functions


The estimation is made in a simple way by means of
the ParamFreqModel function, which is applied to a model
valid VAR(p). In tutorial 3, a VAR(12) model was estimated
corresponding to the tilt interval. We will save your expression
in the frequency domain in the freq_model object:

```{r Estimate freq model 1, eval = TRUE}
freq_model_pre <- ParamFreqModel(model_pre, A0 = FALSE)
```


By specifying that we don't want to use the A0 argument,
we estimate transfer functions that only have
account for interactions between past values of
the variables, regardless of their instantaneous interactions
no lag. To collect these interactions without lag in functions
of transfer, we will do the following:

```{r Estimate freq model 2, eval = TRUE}
freq_model_pre_A0 <- ParamFreqModel(model_pre)
good_model <- freq_model_pre_A0
```


This allows us to calculate the transfer functions. He
next step will be to observe its behavior graphically
and calculate representative indices of these.


## PART 5: Plotting the estimates


To study the graphical representation of functions
of transfer, we will use the PlotTransferFun function.
We will apply it to the freq_model object, generated in the
tutorial 4. The function requires two indices, which
mark the variables of the main process, or feedforward.
In this way, index1 represents the variable x or input of the process
main, while index2 represents your variable y or output.

In our case, the main process (feedforward) characterizes
cardiovascular hemodynamic properties: by increasing the frequency
Cardiac, a rise in blood pressure is achieved. said process
is regulated by a controller process (feedback) that allows
the regulation and control of this heart rate according to the
changes in blood pressure, thus allowing the heart rate to be regulated
according to physiological needs.

To study the function of the main process, that is, the changes in the
blood pressure due to changes in heart rate, we use the
indices index1 = 1 and index2 = 2:

```{r Plot 1, eval = TRUE, fig.align = "center", fig.width = 7}
PlotTransferFun(good_model, 1, 2)
```

To study the controller process, that is, the regulation of the frequency
cardiac from changes in blood pressure, we use the indices
index1 = 2 and index2 = 1:

```{r Plot 2, eval = TRUE, fig.align = "center", fig.width = 7}
PlotTransferFun(good_model, 2, 1)
```


The representations show on the one hand the magnitude of the transfer function
(above) and on the other hand the phase of said function in radians (below). watch yourself now
the changes that occur when taking into account non-lag interactions within the
transfer function of the controlling process compared to
ignore them:

```{r Plot 3, eval = TRUE, fig.align = "center", fig.width = 7}
PlotTransferFun(good_model, 2, 1) # We take these effects into account
PlotTransferFun(freq_model_pre, 2, 1) # We don't take them into account
```

Alternatively, you can perform the graphical representation of the estimate
of both transfer functions assuming open loop conditions.
To do this, we use the open argument. Compare both representations:

```{r Plot 4, eval = TRUE, fig.align = "center", fig.width = 7}
PlotTransferFun(good_model, 2, 1) # We assume closed loop conditions
PlotTransferFun(good_model, 2, 1, open = TRUE) # We assume open loop conditions
```


# PART 6: Calculation of indices


Using the GetExpectedValues function we can obtain
representative indices of each transfer function
for HF and LF bands:
```{r Get expected values, eval = TRUE}
coh <- CalculateCoherence(good_model, 1, 2)
GetExpectedValues(good_model, coherence = coh)
```

To locate a specific value, we will inactivate the argument
str. For example, to locate the representative index
of the controlling process transfer function (feedback)
in the LF band, we use the following expression:
```{r Get expected values 2, eval = TRUE}
GetExpectedValues(good_model, coherence = coh, str = FALSE, weight = FALSE)$LF$Transfer_Functions[2,1,]
```

Now, we will calculate the representative index of the feedforward function
for this same band, LF:
```{r Get expected values 3, eval = TRUE}
GetExpectedValues(good_model, coherence = coh, str = FALSE, weight = FALSE)$LF$Transfer_Functions[1,2,]
```

To obtain the same results in the HF band, it is enough to use the
same expressions, substituting in them LF by HF.

We can also produce estimates by calculating a weighted mean instead of using the coherence:
```{r Get expected values 4, eval = TRUE}
GetExpectedValues(good_model, use.coh = FALSE, weight = TRUE)
```

And we can also produce estimates at the maximum coherence level:
```{r Get expected values 5, eval = TRUE}
GetEstimateAtMaxCoh(good_model, coherence = coh)
```






